\section{Preliminaries}

\subsection{Software Version}

\begin{verbatim}
  BlueSNP 0.1.0
\end{verbatim}

\subsection{Installation}

\begin{itemize}
\item This document picks-up where the installation instructions in the BlueSNP manual leave-off. 
\item \textbf{Before proceeding you must install the BlueSNP package and it's dependencies.} \item See the \textbf{BlueSNP Manual} for installation instructions.
\end{itemize}

\subsection{Download the Tutorial sample data}

Log into the Rhipe VM running on port 2222, user:username, pass:password
\begin{verbatim}
  $ ssh -p 2222 localhost -l username
\end{verbatim}

Download the tutorial materials using wget 
\begin{verbatim}
  $ wget https://github.com/downloads/ibm-bioinformatics/BlueSNP/tutorial_data_v1.tgz
\end{verbatim}

Unzip
\begin{verbatim}
  $ tar xvzf tutorial_data.tgz
\end{verbatim}

\subsection{Copy the tutorial files to the Hadoop filesystem}

Verify that the HDFS directory /user/username exists. (This should have been created at the time of BlueSNP package installation and testing.)

\begin{verbatim}
  $ hadoop fs -ls /user/username
\end{verbatim}

If if does not exist, create it.
\begin{verbatim}
  $ hadoop fs -mkdir /user/username
\end{verbatim}

Copy the tutorial directory from the local filesystem to the HDFS.
\begin{verbatim}
  $ hadoop fs -copyFromLocal tutorial /user/username
\end{verbatim}

Verify that the -copyFromLocal command worked.
\begin{verbatim}
  $ hadoop fs -ls .  # note dot at end
  $ hadoop fs -ls tutorial
\end{verbatim}

Note that a dot (period) is a Hadoop shorthand for the ``current'' HDFS directory, /user/username.

Change to the directory containing the R code.
\begin{verbatim}
  $ cd tutorial/R
\end{verbatim}

Using an additional login window, you may open the tutorial code in a text editor to follow along, or simply cut and paste from this document.

\begin{verbatim}
  $ pico tutorial.R
\end{verbatim}

Start R
\begin{verbatim}
  $ R
\end{verbatim}

